---
title: 'ISYE-6501 : Spring 2020 HW4'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question - 7.1

Exponential smoothing is a technique for smoothing the time series data. It can be used for forecasting the future valuse using the weighted average of the historical time series data. 

As an example, we can use this technique for forecasting the stock prices. I would expect the alpha value to be closer to the 0. I think, stock prices follow certain trend and the prices can be forecasted based on previous observations. So we can use alpha closer to 0 , giving more weightage to the previous observations.


## Question - 7.2

Load libraries required: 

```{r, echo=TRUE}
library(readr)
library(outliers)
library(data.table)
library(xlsx)
```

#### Clean up the envrionment


```{r, echo=TRUE}

v <- c("temperatures","temperatures_vector", "temps_vector_ts") 

rm(ls = v)

```


#### Load the data : 

Load the temperatures data set


```{r, echo=FALSE}

temperatures <- read.table("C:/OMSA GATech/HW4/data 7.2/temps.txt", stringsAsFactors = FALSE, header = TRUE)

```

#### Explore the data

```{r, echo=TRUE}

head(temperatures)

```


Convert temperatures into a time series vector

```{r, echo=TRUE}

temperatures_vector <- as.vector(unlist(temperatures[,2:21]))

temps_vector_ts <- ts(temperatures_vector, start=1996, frequency = 123)

```

Plot the time series temperatures data

```{r, echo=TRUE}

plot(temps_vector_ts)

```

#### Plot the data using Decomposition by moving averages.

```{r, echo=TRUE}

plot(decompose(temps_vector_ts))

```


####  What is exponential smoothing ? 

Exponential smoothing is techique for smoothing time series data. The smoothing functions is represented by the following formula: 

{\displaystyle s_{t}=\alpha x_{t}+(1-\alpha )s_{t-1}=s_{t-1}+\alpha (x_{t}-s_{t-1})}.

where \alpha is smoothing factor between 0 < \alpha < 1. 


Let's use Holt Winters function to apply exponential smoothing on the temperatures data.


```{r, echo=TRUE}

temps_vector_ts_hw <- HoltWinters(temps_vector_ts, alpha = NULL, beta = NULL, gamma = NULL, seasonal = "additive")

```

Summary of the model: 

```{r, echo=TRUE}

summary(temps_vector_ts_hw)

```


Explore the fitted values and plot. 


```{r, echo=TRUE}

head(temps_vector_ts_hw$fitted)

plot(temps_vector_ts_hw)

```


Now export the fitted (smoothed) temperature values from the Holt-Winters function and apply the CUSUM model for analysis.

To detect the end of summer using the given temperature dataset with CUSUM algortihm, following approach has been followed: 

1.  Computed the mu as average of temperature between July 1st and September 30th. 

2. Constant C was computed as 0.5 * standard deviation of the temperature for a given year. 

3. Threshold T was computed as 5 * standard deviation.

4. Then for each year computed the CUSUM using = Max (0, S(t-1) + (mu - x(i) -C)). As we are looking to detect the decreasing temperatures, used mu - x(i). 

5. Based on the analysis, duration wise , 2005 summer was longest and the avg temperature wise, 2010 recorded highest avg temperature. Please refer to the Excel spreadsheets for the CUSUM computations, analysis and charts.

6. Spreadsheet details :
  
   a. Temps sheet has the fitted data from the Holt-Winters model. 1996 data will not be present as that was used by HW for fitting the model.
   
   b. Cusum Analysis-1 sheet has CUSUM calculations by each year. Then highlighted the cooling off date based on CUSUM change detection for each column. 
   
   c. Cusum Analysis-2 sheet has analysis for interpreting the duration of the summer and avg summer temperature charts.
   
   

#### Resources: 

1. BlueJeans office hour video : https://bluejeans.com/playback/s/KmWjgZ8lfIoMtoJML94yuazOM7yvY8NQxAhUZjTtZH9663EjnYklpV5vCa4VdCTH
2. https://en.wikipedia.org/wiki/Exponential_smoothing

